Flattering Advice: Avoiding Disappointment in Advice-Giving

### 2025-01-29 manuscript/manuscript.qmd
Add in "interpersonal concerns"

People go to great lengths to avoid unpleasant information,
People often go to great lengths to avoid unpleasant information,
deriving belief-based utility from maintaining favorable views of
themselves [@Golman2017; @Ho2018; @Loewenstein2018]. Moreover, they
often prefer advisers who withhold bad news from them [@Shalvi2019] and
punish those who communicate bad news [@John2019]. Advisers,
anticipating these preferences and consequences face a dilemma: Should
they provide the accurate but potentially painful truth, or shade their
recommendations to preserve the advisee's mood and the adviser's own
standing? Advisers who fear they will be blamed or disliked for
undermining someone's confidence may reasonably avoid candid feedback,
particularly if they don't incur a cost when the advisee makes a
mistake.

>>>>

While much of the literature on advice assumes that advisers aim to provide
the most helpful advice, thier motivations are often more complex [@Yaniv2012;@Dalal2010; @Rader2017]. 
Advisers are not only concerned with the accuracy and usefulness of their advice 
but also with their relationships with advisees. Interpersonal concerns—such
as maintaining a positive image, avoiding conflict, or preserving social 
harmony—can shape the way advice is delivered [@Van2009; @Schaerer2018; @Chentsova2012].
In particular, advisers may strategically soften their messages or withhold 
unfavorable information to manage how they are perceived or to prevent discomfort
for both themselves and the advisee [@Sniezek2001;@Levine2020]. These interpersonal dynamics 
suggest that advice is not always optimized for the recipient’s best interest but
is instead filtered through social and psychological considerations.

Just as advisers may hesitate to provide unwelcome truths, advisees themselves are o
ften resistant to hearing them. People go to great lengths to avoid unpleasant information,
People often go to great lengths to avoid unpleasant information,
deriving belief-based utility from maintaining favorable views of
themselves [@Golman2017; @Ho2018; @Loewenstein2018]. Moreover, they
often prefer advisers who withhold bad news from them [@Shalvi2019] and
punish those who communicate bad news [@John2019]. Advisers,
anticipating these preferences and consequences face a dilemma: Should
they provide the accurate but potentially painful truth, or shade their
recommendations to preserve the advisee's mood and the adviser's own
standing? Advisers who fear they will be blamed or disliked for
undermining someone's confidence may reasonably avoid candid feedback,
particularly if they don't incur a cost when the advisee makes a
mistake.

### 2025-01-29 manuscript/manuscript.tex
Add in "interpersonal concerns"

People go to great lengths to avoid unpleasant information, People often
go to great lengths to avoid unpleasant information, deriving
belief-based utility from maintaining favorable views of themselves
(Golman et al., 2017; Ho et al., 2021; Loewenstein \& Molnar, 2018).
Moreover, they often prefer advisers who withhold bad news from them
(Shalvi et al., 2019) and punish those who communicate bad news (John et
al., 2019). Advisers, anticipating these preferences and consequences
face a dilemma: Should they provide the accurate but potentially painful
truth, or shade their recommendations to preserve the advisee's mood and
the adviser's own standing? Advisers who fear they will be blamed or
disliked for undermining someone's confidence may reasonably avoid
candid feedback, particularly if they don't incur a cost when the
advisee makes a mistake.

@@

@@

@@

@@

@@

>>>>

While much of the literature on advice assumes that advisers aim to
provide the most helpful advice, thier motivations are often more
complex (Dalal \& Bonaccio, 2010; Rader et al., 2017; Yaniv \&
Choshen-Hillel, 2012). Advisers are not only concerned with the accuracy
and usefulness of their advice but also with their relationships with
advisees. Interpersonal concerns---such as maintaining a positive image,
avoiding conflict, or preserving social harmony---can shape the way
advice is delivered (Chentsova-Dutton \& Vaughn, 2012; Schaerer et al.,
2018; Van Swol, 2009). In particular, advisers may strategically soften
their messages or withhold unfavorable information to manage how they
are perceived or to prevent discomfort for both themselves and the
advisee (Levine et al., 2020; Sniezek \& Van Swol, 2001). These
interpersonal dynamics suggest that advice is not always optimized for
the recipient's best interest but is instead filtered through social and
psychological considerations.

Just as advisers may hesitate to provide unwelcome truths, advisees
themselves are o ften resistant to hearing them. People go to great
lengths to avoid unpleasant information, People often go to great
lengths to avoid unpleasant information, deriving belief-based utility
from maintaining favorable views of themselves (Golman et al., 2017; Ho
et al., 2021; Loewenstein \& Molnar, 2018). Moreover, they often prefer
advisers who withhold bad news from them (Shalvi et al., 2019) and
punish those who communicate bad news (John et al., 2019). Advisers,
anticipating these preferences and consequences face a dilemma: Should
they provide the accurate but potentially painful truth, or shade their
recommendations to preserve the advisee's mood and the adviser's own
standing? Advisers who fear they will be blamed or disliked for
undermining someone's confidence may reasonably avoid candid feedback,
particularly if they don't incur a cost when the advisee makes a
mistake.

@@

\bibitem[\citeproctext]{ref-Chentsova2012}
Chentsova-Dutton, Y. E., \& Vaughn, A. (2012). Let me tell you what to
do: Cultural differences in advice-giving. \emph{Journal of
Cross-Cultural Psychology}, \emph{43}(5), 687--703.

\bibitem[\citeproctext]{ref-Dalal2010}
Dalal, R. S., \& Bonaccio, S. (2010). What types of advice do
decision-makers prefer? \emph{Organizational Behavior and Human Decision
Processes}, \emph{112}(1), 11--23.

@@

\bibitem[\citeproctext]{ref-Levine2020}
Levine, E. E., Roberts, A. R., \& Cohen, T. R. (2020). Difficult
conversations: Navigating the tension between honesty and benevolence.
\emph{Current Opinion in Psychology}, \emph{31}, 38--43.

@@

\bibitem[\citeproctext]{ref-Rader2017}
Rader, C. A., Larrick, R. P., \& Soll, J. B. (2017). Advice as a form of
social influence: Informational motives and the consequences for
accuracy. \emph{Social and Personality Psychology Compass},
\emph{11}(8), e12329.

\bibitem[\citeproctext]{ref-Schaerer2018}
Schaerer, M., Tost, L. P., Huang, L., Gino, F., \& Larrick, R. (2018).
Advice giving: A subtle pathway to power. \emph{Personality and Social
Psychology Bulletin}, \emph{44}(5), 746--761.

@@

\bibitem[\citeproctext]{ref-Sniezek2001}
Sniezek, J. A., \& Van Swol, L. M. (2001). Trust, confidence, and
expertise in a judge-advisor system. \emph{Organizational Behavior and
Human Decision Processes}, \emph{84}(2), 288--307.

@@

\bibitem[\citeproctext]{ref-Van2009}
Van Swol, L. M. (2009). The effects of confidence and advisor motives on
advice utilization. \emph{Communication Research}, \emph{36}(6),
857--873.

\bibitem[\citeproctext]{ref-Yaniv2012}
Yaniv, I., \& Choshen-Hillel, S. (2012). Exploiting the wisdom of others
to make better decisions: Suspending judgment reduces egocentrism and
increases accuracy. \emph{Journal of Behavioral Decision Making},
\emph{25}(5), 427--434.
